---
title: "Data Gathering"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(janitor)
library(tidycensus)
library(leaflet)
library(lubridate)
library(stringr)
library(rstanarm)
library(naniar)
```

Dataset
```{r}
acled_data <- read_xlsx("raw_data/protest_data.xlsx") %>%
  clean_names() %>%
  mutate(assoc_actor_1 = as.factor(assoc_actor_1),
         event_type = as.factor(event_type),
         sub_event_type = as.factor(sub_event_type)) %>%
  filter(!event_type == "Strategic developments") %>%
  mutate(event_date = ymd(event_date))
```

General Protest Info
```{r}
protesters_type <- levels(acled_data$assoc_actor_1)
event_type <- levels(acled_data$event_type)
sub_event_type <- levels(acled_data$sub_event_type)

blm_terms <- c("BLM", "NAACP", "African American", "African", "Black")
labor_terms <- c("Federation", "Union", "Labor", "Labour", "Worker", "AFL")
antifa_terms <- c("Antifa", "antifa", "Anti Facist")
militia_terms <- c("Militia", "Patriot", "Proud Boy", "Boogaloo", "KKK", 
                   "QAnon")
pro_police_terms <- c("Pro-Police", "Pro Police", "Blue")
republican_terms <- c("Republican", "GOP")
democrat_terms <- c("Dem", "Democrat")


acled_data_new <- acled_data %>%
  select(-c(iso, event_id_cnty, event_id_no_cnty, year, time_precision, region, 
            country, admin3, geo_precision, source_scale, source)) %>%
  mutate(assoc_actor_1 = as.character(assoc_actor_1)) %>%
  mutate(blm = ifelse(str_detect(assoc_actor_1, 
                                 paste(blm_terms, collapse = '|')), 
                      yes = TRUE, no = FALSE)) %>%
  mutate(labor = ifelse(str_detect(assoc_actor_1, 
                                   paste(labor_terms, collapse = '|')), 
                      yes = TRUE, no = FALSE)) %>%
  mutate(antifa = ifelse(str_detect(assoc_actor_1, 
                                   paste(antifa_terms, collapse = '|')), 
                      yes = TRUE, no = FALSE)) %>%
  mutate(militia = ifelse(str_detect(assoc_actor_1, 
                                   paste(militia_terms, collapse = '|')), 
                      yes = TRUE, no = FALSE)) %>%
  mutate(pro_police = ifelse(str_detect(assoc_actor_1, 
                                   paste(pro_police_terms, collapse = '|')), 
                      yes = TRUE, no = FALSE)) %>%
  mutate(republican = ifelse(str_detect(assoc_actor_1, 
                                   paste(republican_terms, collapse = '|')), 
                      yes = TRUE, no = FALSE)) %>%
  mutate(democrat = ifelse(str_detect(assoc_actor_1, 
                                   paste(democrat_terms, collapse = '|')), 
                      yes = TRUE, no = FALSE)) %>%
  mutate(unknown = ifelse(is.na(assoc_actor_1), yes = TRUE, no = FALSE)) %>%
  mutate(all_actors = TRUE,
         all_violence = TRUE) %>%
  rename("district" = admin2) %>%
  transform(district = paste(district, admin1, sep=", ")) %>%
  pivot_longer(cols = blm:democrat,
               names_to = "subject",
               values_to = "group_boolean")

```

Mapping
```{r}
census_api_key("c272499f01a005c075c74055db3526e316e1836d")

rural <- get_decennial(geography = "state",
                       variables = c("P001001", "P002005"),
                       year = 2010,
                       output = "wide",
                       geometry = TRUE) %>%
  rename(state = NAME) %>%
  mutate(prop_rural = P002005/P001001,
         state = reorder(state, prop_rural))

subset <- acled_data_new %>%
  mutate(event_date = ymd(event_date)) %>%
  filter(event_date <= "2020-06-01")

pal <- 
   colorFactor(palette = "Reds",
               levels = c("Peaceful protest", "Protest with intervention", 
                          "Violent demonstration", "Mob violence"))

#leaflet(rural) %>%
#  addTiles() %>%
#  addCircleMarkers(lng = subset$longitude, lat = subset$latitude, radius = 0.1,
#                   color = ~pal(acled_data_new$sub_event_type))
```


Prepping ACLED data for violence modeling
```{r}
acled_modeling <- acled_data_new %>%
  filter(subject == "blm") %>%
  mutate(violence = ifelse(sub_event_type == "Peaceful protest", 
                           yes = FALSE, no = TRUE))

count_violent <- acled_modeling %>%
  filter(violence == TRUE) %>%  
  group_by(district) %>%
  summarize(number_violent = n(), .groups = "drop")

count_protests <- acled_modeling %>%
  group_by(district) %>%
  summarize(number_protests = n(), .groups = "drop")

model_violence <- full_join(count_violent, count_protests, by = "district")

model_violence[is.na(model_violence)] <- 0
  
model_violence <- model_violence %>%
  mutate(proportion_violent = number_violent / number_protests,
         violence = ifelse(proportion_violent == 0,
                                      yes = FALSE,
                                      no = TRUE)) %>%
  select(!violence)
```

Function for inputting new parameters from Census data
```{r}
district_formatting <- function(ds){
  ds_clean <- ds[-1 ,] %>%
    rename("district" = NAME) %>%
    separate(col = district, sep = ", ", into = c("district", "state"))
  correct_district <- ds_clean %>%
    mutate(district = sub(ds_clean$district, pattern = " [[:alpha:]]*$", 
                        replacement = "")) %>%
    transform(district = paste(district, state, sep=", ")) %>%
    select(!state)
  return(correct_district)
  }
```

Reading in parameters
```{r}
# gini
gini <- read_csv("raw_data/census_gini.csv", col_types = cols(
  GEO_ID = col_character(),
  NAME = col_character(),
  B19083_001E = col_character(),
  B19083_001M = col_character()
))
gini_final <- district_formatting(gini) %>%
  select(B19083_001E, district) %>%
  rename("gini_score" = B19083_001E) %>%
  mutate(gini_score = as.numeric(gini_score))

# poverty
poverty <- read_csv("raw_data/census_poverty.csv", col_types = cols(
  .default = col_character()
)) %>%
  select(S1701_C03_001E, NAME)
poverty_final <- district_formatting(poverty) %>%
  rename("poverty" = S1701_C03_001E) %>%
  mutate(poverty = as.numeric(poverty))

# income
income <- read_csv("raw_data/census_income.csv", col_types = cols(
  .default = col_character()
)) %>%
  select(NAME, S1901_C01_012E)
income_final <- district_formatting(income) %>%
  rename("income" = S1901_C01_012E) %>%
  mutate(income = as.numeric(income))

# population
population <- read_csv("raw_data/census_population.csv", col_types = cols(
  .default = col_character()
)) %>%
  select(NAME, B01003_001E)
population_final <- district_formatting(population) %>%
  rename("population" = B01003_001E) %>%
  mutate(population = as.numeric(population))

# unemployment
unemployment <- read_csv("raw_data/census_unemployment.csv", col_types = cols(
  .default = col_character()
))
unemployment_final <- district_formatting(unemployment) %>%
  select(district, S2301_C04_001E) %>%
  rename("unemployment" = S2301_C04_001E) %>%
  replace_with_na(replace = list(unemployment = "N")) %>%
  mutate(unemployment = as.numeric(unemployment))

# education
education_final <- district_formatting(unemployment) %>%
  select(S2301_C01_031E, S2301_C01_032E, S2301_C01_033E, S2301_C01_034E, 
         S2301_C01_035E, district) %>%
  rename("pop" = S2301_C01_031E,
         "less_than_hs" = S2301_C01_032E,
         "hs" = S2301_C01_033E,
         "some_college" = S2301_C01_034E,
         "college" = S2301_C01_035E) %>%
  replace_with_na(replace = list(pop = "N")) %>%
  replace_with_na(replace = list(less_than_hs = "N")) %>%
  mutate(less_than_hs = as.numeric(less_than_hs),
         pop = as.numeric(pop),
         prop_less_than_hs = less_than_hs / pop) %>%
  select(prop_less_than_hs, district)

# police killings
states <- tibble(state_name = state.name) %>%
  bind_cols(tibble(state = state.abb))
police_killings_final <- read_excel("raw_data/police_killings.xlsx") %>%
  left_join(states, by = "state") %>%
  select(!state) %>%
  rename("state" = state_name,
         "district" = county) %>%
  drop_na() %>%
  transform(district = paste(district, state, sep=", ")) %>%
  group_by(district) %>%
  summarize(police_killings = n(), .groups = "drop")

# ideology
ideology <- read_csv("raw_data/ideology.csv", col_types = cols(
  county_fips = col_double(),
  mrp_ideology_mean = col_double(),
  mrp_sd = col_double(),
  mrp_lower = col_double(),
  mrp_upper = col_double(),
  raw_ideology_mean = col_double(),
  sample_size = col_double(),
  abb = col_character(),
  state_fips = col_double(),
  county_name = col_character(),
  population_2010 = col_character(),
  state = col_character(),
  obama_pc_t2012 = col_double(),
  romney_pct_2012 = col_character(),
  other_pct_2012 = col_character(),
  obama_pct_2008 = col_double(),
  mccain_pct_2008 = col_character(),
  other_pct_2008 = col_character()
)) %>%
  select(mrp_ideology_mean, county_name) %>%
  rename("NAME" = county_name,
         "ideology" = mrp_ideology_mean)
ideology_final <- district_formatting(ideology)

# mobility
mobility <- read_csv("raw_data/mobility.csv", col_types = cols(
  cty = col_character(),
  Name = col_character(),
  Household_Income_rP_gP_p25 = col_double()
)) %>%
  rename("district" = Name) %>%
  separate(col = district, sep = ", ", into = c("district", "state")) %>%
  left_join(states, by = "state") %>%
  select(!state) %>%
  rename("state" = state_name)

mobility_final <- mobility %>%
  mutate(district = sub(mobility$district, pattern = " [[:alpha:]]*$", 
                        replacement = "")) %>%
  transform(district = paste(district, state, sep=", ")) %>%
  select(Household_Income_rP_gP_p25, district) %>%
  rename("mobility" = Household_Income_rP_gP_p25)

```

County Health Data
```{r}
health_data <- read_excel("raw_data/ranked_measure_data.xlsx", 
                                  skip = 1) %>%
  clean_names() %>%
  drop_na(county) %>%
  mutate(district = sub(county, pattern = " [[:alpha:]]*$", 
                        replacement = "")) %>%
  transform(district = paste(district, state, sep = ", ")) %>%
  rename("premature_death_rate" = years_of_potential_life_lost_rate,
         "housing_problems" = percent_severe_housing_problems)

health_final <- health_data %>%
  select(district, premature_death_rate, teen_birth_rate, 
         preventable_hospitalization_rate, high_school_graduation_rate, 
         violent_crime_rate, housing_problems)

health_cont_data <- read_excel("raw_data/health_cont.xlsx", 
                                  skip = 1) %>%
  clean_names() %>%
  rename("physical_distress" = percent_frequent_physical_distress,
         "mental_distress" = percent_frequent_mental_distress) %>%
  drop_na(county) %>%
  mutate(district = sub(county, pattern = " [[:alpha:]]*$", 
                        replacement = "")) %>%
  transform(district = paste(district, state, sep = ", "))

health_cont_final <- health_cont_data %>%
  select(district, physical_distress, mental_distress, life_expectancy, 
         segregation_index, segregation_index_2, percent_homeowners, 
         percent_black, percent_non_hispanic_white, percent_hispanic, 
         homicide_rate, firearm_fatalities_rate)

model_health <- full_join(x = health_final, y = health_cont_final, 
                          by = "district") %>%
  full_join(model_violence, by = "district")

glimpse(model_health)
```

```{r}
# teen_birth_rate, high_school_graduation_rate, violent_crime_rate, physical_distress, segregation_index_2, life_expectancy, percent_homeowners, percent_black, number_protests

fit_health <- stan_glm(data = model_health,
                formula = number_violent ~ number_protests, 
                refresh = 0)
                
print(fit_health, digits = 5)
```


Parameter dataset

```{r}
model <- full_join(x = gini_final, y = poverty_final, by = "district") %>%
  full_join(income_final, by = "district") %>%
  full_join(population_final, by = "district") %>%
  full_join(education_final, by = "district") %>%
  full_join(police_killings_final, by = "district") %>%
  full_join(ideology_final, by = "district") %>%
  full_join(mobility_final, by = "district") %>%
  full_join(health_final, by = "district") %>%
  full_join(health_cont_final, by = "district") %>%
  full_join(model_violence, by = "district") %>%
  select(!c(mental_distress, segregation_index, mobility, premature_death_rate, 
            preventable_hospitalization_rate, homicide_rate, 
            firearm_fatalities_rate, proportion_violent))

glimpse(model)
```

### All Parameters
gini_score
poverty
income
population
education


Modeling: all
```{r}
fit <- stan_glm(data = model,
                formula = number_violent ~ gini_score + prop_less_than_hs + 
                  police_killings + ideology + teen_birth_rate + high_school_graduation_rate + 
                  violent_crime_rate + housing_problems + physical_distress + 
                  life_expectancy + segregation_index_2 + percent_homeowners + 
                  number_protests, 
                refresh = 0)
                
print(fit, digits = 5)
```



Modeling: Income Inequality
```{r}
model_violence_gini <- model_violence %>%
  left_join(gini_final, by = "district") %>%
  select(district, violence, gini_score, proportion_violent)

fit_obj_gini <- stan_glm(data = model_violence_gini,
                    proportion_violent ~ gini_score - 1,
                    refresh = 0)

new_obs_gini <- tibble(gini_score = c(0.3288, 0.4238, 0.4489, 0.6592))

pp_gini <- posterior_predict(fit_obj_gini, newdata = new_obs_gini) %>%
  as_tibble() %>%
  mutate_all(as.numeric) %>%
  rename("min" = `1`,
         "low" = `2`,
         "high" = `3`,
         "max" = `4`) %>%
  pivot_longer(cols = min:max,
               names_to = "gini",
               values_to = "proportion_violent")

pp_gini_mean <- pp_gini %>%
  group_by(gini) %>%
  summarize(mean_violence = mean(proportion_violent), .groups = "drop") %>%
  arrange(mean_violence)

pp_gini_mean$mean_violence

ggplot(pp_gini, aes(x = proportion_violent, fill = gini)) +
  geom_histogram(aes(y = after_stat(count/sum(count))), bins = 75,
                 color = "white", position = "identity", alpha = 0.5) +
  geom_vline(xintercept = pp_gini_mean$mean_violence[1]) +
  geom_vline(xintercept = pp_gini_mean$mean_violence[2]) +
  geom_vline(xintercept = pp_gini_mean$mean_violence[3]) +
  geom_vline(xintercept = pp_gini_mean$mean_violence[4]) +
  scale_y_continuous(labels = scales::percent_format()) + 
  labs(title = "Posterior Distribution for Violence based on Gini",
       x = "Proportion Violent",
       y = "Probability") +
  theme_bw()

```

Modeling: Poverty
```{r}
model_violence_poverty <- model_violence %>%
  left_join(poverty_final, by = "district") %>%
  select(district, violence, poverty, proportion_violent) %>%
  drop_na(poverty)

fit_obj_poverty <- stan_glm(data = model_violence_poverty,
                    proportion_violent ~ poverty - 1,
                    refresh = 0)

quantile(model_violence_poverty$poverty)

new_obs_poverty <- tibble(poverty = c(8.7, 15.5))

pp_poverty <- posterior_predict(fit_obj_poverty, newdata = new_obs_poverty) %>%
  as_tibble() %>%
  mutate_all(as.numeric) %>%
  rename("low" = `1`,
         "high" = `2`) %>%
  pivot_longer(cols = low:high,
               names_to = "poverty",
               values_to = "proportion_violent")

ggplot(pp_poverty, aes(x = proportion_violent, fill = poverty)) +
  geom_histogram(aes(y = after_stat(count/sum(count))), bins = 75,
                 color = "white", position = "identity", alpha = 0.5) +
  scale_y_continuous(labels = scales::percent_format()) + 
  labs(title = "Posterior Distribution for Poverty Rate",
       x = "Poverty Rate",
       y = "Probability") +
  theme_bw()

```

Modeling: Income
```{r}
model_violence_income <- model_violence %>%
  left_join(income_final, by = "district") %>%
  select(district, violence, income, proportion_violent) %>%
  drop_na(income)

fit_obj_income <- stan_glm(data = model_violence_income,
                    proportion_violent ~ income - 1,
                    refresh = 0)

new_obs_income <- tibble(income = c(53518, 72736))

pp_income <- posterior_predict(fit_obj_income, newdata = new_obs_income) %>%
  as_tibble() %>%
  mutate_all(as.numeric) %>%
  rename("low" = `1`,
         "high" = `2`) %>%
  pivot_longer(cols = low:high,
               names_to = "income",
               values_to = "proportion_violent")

ggplot(pp_income, aes(x = proportion_violent, fill = income)) +
  geom_histogram(aes(y = after_stat(count/sum(count))), bins = 75,
                 color = "white", position = "identity", alpha = 0.5) +
  scale_y_continuous(labels = scales::percent_format()) + 
  labs(title = "Posterior Distribution for Violence based on Income",
       x = "Proportion Violent",
       y = "Probability") +
  theme_bw()

```






from meeting w Mitchell — what we'll end up doing is basically this

```{r}
new_obssss <- tibble(gini_score = input$whatever,
                  poverty = input$whatever)

posterior_predict(dataset, new_obssss)
```

interesting final projects:
https://nidal-morrison.shinyapps.io/urgent_fury/
https://wyatthurt.shinyapps.io/water_conflict/
